---
title: "Part IV"
author: "Krishna Kumar Shrestha"
date: "10/31/2020"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)


```

# Chapter 17:SAMPLING DISTRIBUTIONS AND CONFIDENCE

Exercise 17.1

teacher wants to test all of the 10th-grade students at his school to
gauge their basic mathematical understanding, but the photocopier breaks
after making only six copies of the test. With no other choice, he
chooses six students at random to take the test. Their results, recorded
as a score out of 65, have a sample mean of 41.1. The standard deviation
of the marks of this test is known to be 11.3. a. Find the standard
error associated with the mean test score.

```{r}
mean <- 41.1
sd <- 11.3
n <- 6
se <- sd / sqrt(n)
```

b.  Assuming the scores themselves are normally distributed, evaluate
    the probability that the mean score lies between 45 and 55 if the
    teacher took another sample of the same size.

```{r}
pnorm(55, mean, se) - pnorm(45, mean, se)
```

c.  A student who gets less than half the questions correct receives a
    failing grade (F). Find the probability that the average score is an
    F based on another sample of the same size.

```{r}
pnorm(65 / 2, mean, se)
```

A marketing company wants to find out which of two energy drinks
teenagers prefer---drink A or drink B. It surveys 140 teens, and the
results indicate that only 35 percent prefer drink A. d. Use a quick
check to decide whether it is valid to use the normal distribution to
represent the sampling distribution of this proportion.

```{r}
n <- 140
p <- 0.35 #probability of A being choosed
p * n >= 5 & n * (1 - p) >= 5





```

e.  What is the probability that in another sample of the same size, the
    proportion of teenagers who prefer drink A is greater than 0:4?

```{r}

se <- sqrt(p * (1 - p) / n)   # to be able to use it in the normal formula for
                              # sample below
1 - pnorm(0.4, p, se)

```

f.  Find the two values of this sampling distribution that identify the
    central 80 percent of values of the proportion of interest.

```{r}
upper<-qnorm(0.9,p,se)
lower<-qnorm(0.1,p,se)
```

In Section 16.2.4, the time between cars passing an individual's
location was modeled using an exponential distribution. Say that on the
other side of town, her friend is curious about a similar problem.
Standing outside her house, she records 63 individual times between cars
passing. These sampled times have a mean of ¯x = 37:8 seconds with a
standard deviation of s = 34:51 seconds.

g.  The friend inspects a histogram of her raw measurements and notices
    that her raw data are heavily right-skewed. Briefly identify and
    describe the nature of the sampling distribution with respect to the
    sample mean and calculate the appropriate standard error.

```{r}
n<-63
xmean<-37.8
s<-34.51



# This is a normal curve as it is regarding a mean of samples.
se <- s/sqrt(n)
df <- n - 1
```

h.  Using the standard error from (g) and the appropriate probability
    distribution, calculate the probability that in another sample of
    the same size, the sample mean time between cars passing is as
    follows:
i.  More than 40 seconds

```{r}
#i but as it is regarding a sample, it works as a normal.
#all 

newx <- (40 - xmean) / se
1 - pt(newx, df)

```

ii. Less than half a minute

```{r}
newx <- (30 - xmean) / se
pt(newx, df)
```

iii. Between the given sample mean and 40 seconds

```{r}
newx <- (40 - xmean) / se
pt(newx,df)-0.5
```

Exercise 17.2

casual runner records the average time it takes him to sprint 100
meters. He completes the dash 34 times under identical conditions and
finds that the mean of these is 14.22 seconds. Assume that he knows the
standard deviation of his runs is X = 2:9 seconds. a. Construct and
interpret a 90 percent confidence interval for the true mean time.

```{r}
n=34
x.bar<-14.22
sigma<-2.9

alpha<-1-.9

x.bar+c(-1,1)*qnorm(1-alpha/2)*(sigma/sqrt(n))
```

b.  Repeat (a), but this time, assume that the standard deviation is not
    known and that s = 2:9 is estimated from the sample. How, if at all,
    does this change the interval?

```{r}
s<-2.9
x.bar+c(-1,1)*qt(1-alpha/2,df=n-1)*(s/sqrt(n))
```

In a particular country, the true proportion of citizens who are left
handed or ambidextrous is unknown. A random sample of 400 people is
taken, and each individual is asked to identify with one of three
options: right-handed only, left-handed only, or ambidextrous. The
results show that 37 selected left-handed and 11 selected ambidextrous.
c. Calculate and interpret a 99 percent CI for the true proportion of
left-handed-only citizens.

```{r}
n<-400
# Option A is right-handed
# Option B is left-handed
# Option C is ambidextrous

p.b<-37/n
p.c<-11/n
p.a<-1-p.b-p.c
alpha<-1-0.99
p.b+
      (c(-1,1)*qnorm(1-alpha/2))*
      sqrt(p.b*(1-p.b)/n)


```

d.  Calculate and interpret a 99 percent CI for the true proportion of
    citizens who are either left-handed or ambidextrous.

```{r}
(p.b + p.c)+
      c(-1, 1) * qnorm(1 - alpha / 2) *
      (sqrt((p.b + p.c) * (1 - (p.b + p.c)) / n))
```

In Section 17.2.4, the technical interpretation of a CI with respect to
its confidence level was described as the proportion of many similar
intervals (that is, when calculated for samples of the same size from
the same population) that contain the true value of the parameter of
interest.

e.  Your task is to write an example to demonstrate this behavior of
    confidence intervals using simulation. To do so, follow these
    instructions: -- Set up a matrix (see Chapter 3) filled with NAs
    (Chapter 6) that has 5,000 rows and 3 columns. -- Use skills from
    Chapter 10 to write a for loop that, at each of 5,000 iterations,
    generates a random sample of size 300 from an exponential
    distribution with rate parameter e = 0:1 (Section 16.2.4). --
    Evaluate the sample mean and sample standard deviation of each
    sample, and use these quantities with the critical values from the
    appropriate sampling distribution to calculate a 95 percent CI for
    the true mean of the distribution. -- Within the for loop, the
    matrix should now be filled, row by row, with your results. The
    first column will contain the lower limit, the second will contain
    the upper limit, and the third column will be a logical value that
    is TRUE if the corresponding interval contains the true mean of 1=e
    and that is FALSE otherwise. -- When the loop is completed, compute
    the proportion of TRUEs in the third column of the filled matrix.
    You should find that this proportion is close to 0:95; this will
    vary randomly each time you rerun the loop.

```{r}
x.mat<-matrix(NA,5000,3)
n<-300
lambda<-0.1
mu<-1/lambda

for(i in 1:nrow(x.mat)) {
      
      sample <- rexp(n, lambda)

      limits <- mean(sample) +
            c(-1, 1) * qt(1-0.05/2, df= n-1) *
            (sd(sample) / sqrt(n))
      x.mat[i, 1:2] <- limits
      x.mat[i,3]<- mu>=limits[1]&mu<=limits[2] #This returns a T F vector

}

mean(x.mat[,3])


#Forth Part ----
#plot the 100 first CI with the average in the middle

xaxis<-seq(mu-3,mu+3,length=100)
yaxis<-1:100

plot(xaxis, yaxis, type = "n") +
      abline(v = 10, lty = 2) +
      for (i in 1:length(xaxis)) {
            lines(c(x.mat[i, 1], x.mat[i, 2]),
                  c(i, i),
                  col = if (x.mat[i, 1] < 10 & x.mat[i, 2] > 10) {
                        "blue"
                  } else{
                        "red"
                  })
            
            
            
      }
dev.off()    
```

# chapter 18 HYPOTHESIS TESTING

Exercise 18.1 a. Adult domestic cats of a certain breed are said to have
an average weight of 3.5 kilograms. A feline enthusiast disagrees and
collects a sample of 73 weights of cats of this breed. From her sample,
she calculates a mean of 3.97 kilograms and a standard deviation of 2.21
kilograms. Perform a hypothesis test to test her claim that the true
mean weight  is not 3.5 kilograms by setting up the appropriate
hypothesis, carrying out the analysis, and interpreting the p-value
(assume the significance level is  = 0:05).

```{r}

mu<-3.5

n<-73
x.bar<-3.97
x.sd<-2.21
#significance = 0.05
#H0<- 3.35
#ha != h0

#solution
t<-(x.bar-mu)/(x.sd/sqrt(n))
p<-pt(-t,df=n-1)+(1-pt(t,df=n-1))
# the probability is greater than the alpha (significance value). For this 
# reason, we don't have reasons to believe the H0 is not true.

CI<- x.bar+c(-1,1)*qt(0.975,n-1)*x.sd/sqrt(n)
#the CI contains the H0. reassuring the t-test
```

b.  Suppose it was previously believed that the mean magnitude of
    seismic events off the coast of Fiji is 4.3 on the Richter scale.
    Use the data in the mag variable of the ready-to-use quakes data
    set, providing 1,000 sampled seismic events in that area, to test
    the claim that the true mean magnitude is in fact greater than 4.3.
    Set up appropriate hypotheses, use t.test (conduct the test at a
    significance level of  = 0:01), and draw a conclusion.

```{r}
#mean magnitude of sismic events on fiji
mu<-4.3
# Hypothesis Test is: ha>h0
alpha<-0.01
remove(x.bar)

t.test(quakes$mag,mu=mu,alternative="greater",conf.level=1-alpha)
#p.value is too small. A strong evidence the true magnitude is greater than 4.3
```

c.  Manually compute a two-sided confidence interval for the true mean
    of (b)

```{r}
n<-length(quakes$mag)
CI<-mean(quakes$mag)+
      c(-1,1)*qt(1-0.01/2,n-1)*
      sd(quakes$mag)/sqrt(n)
#the CI interval does not contain the h0 mean for the true mean value.
```

Exercise 18.2

In the package MASS you'll find the data set anorexia, which contains
data on pre- and post-treatment weights (in pounds) of 72 young women
suffering from the disease, obtained from Hand et al. (1994). One group
of women is the control group (in other words, no intervention), and the
other two groups are the cognitive behavioral program and family support
intervention program groups. Load the library and ensure you can access
the data frame and understand its contents. Let d denote the mean
difference in weight, computed as (post-weight 􀀀 pre-weight).

\<\n\> a. Regardless of which treatment group the participants fall
into, conduct and conclude an appropriate hypothesis test with  = 0:05
for the entire set of weights for the following hypotheses:

```{r}
library(MASS)
data(anorexia)
alpha<-0.05
#H0: d.mu=0 // HA: d.mu>0 The dist is a Student one

t.test(anorexia[,3],
       anorexia[,2],
       alternative = "greater",
       paired=TRUE)

#there is strong evidence to accept H0, as the p.value is small.
```

b.  Next, conduct three separate hypothesis tests using the same defined
    hypotheses, based on which treatment group the participants fall
    into. What do you notice?

```{r}
# Now a test for all the treatments
#Cont
t.test(anorexia[anorexia$Treat=="Cont",3],
       anorexia[anorexia$Treat=="Cont",2],
       alternative="greater",
       paired=TRUE)
# There is no statistical evidence to reject the claim that there is no 
# difference between  pre and post mean in the control group.

#CBT
t.test(anorexia[anorexia$Treat=="CBT",3],
       anorexia[anorexia$Treat=="CBT",2],
       alternative="greater",
       paired=TRUE)
# There is mild statistical evidence to accept that the post mean is greater 
# than the pre mean in the CBT group.

#FT
t.test(anorexia[anorexia$Treat=="FT",3],
       anorexia[anorexia$Treat=="FT",2],
       alternative="greater",
       paired=TRUE)

# There is strong statistical evidence to accept that in the FT group the mean 
#of post measurement is greater than the pre weight.
```

Another ready-to-use data set in R is PlantGrowth (Dobson, 1983), which
records a continuous measure of the yields of a certain plant, looking
at the potential effect of two supplements administered during growth to
increase the yield when compared to a control group with no supplement.

c.  Set up hypotheses to test whether the mean yield for the control
    group is less than the mean yield from a plant given either of the
    treatments. Determine whether this test should proceed using a
    pooled estimate of the variance or whether Welch's t-test would be
    more appropriate.

```{r}
head(PlantGrowth)
levels(PlantGrowth[, 2])
data.1 <- PlantGrowth[PlantGrowth$group == "ctrl", ]
data.2 <- PlantGrowth[PlantGrowth$group != "ctrl", ]
#Hypotheses--> H0: ctrlmean-trtmean=0 // HA: ctrlmean-trtmean<0 

#Rule of thumb
sd(PlantGrowth[PlantGrowth$group != "ctrl",1])/
      sd(PlantGrowth[PlantGrowth$group == "ctrl",1])
      
# The rule of thumb states that there is enought to accept both variances are
# the same for that, we well use the pooled variance estimative test
```

d.  Conduct the test and make a conclusion (assuming normality of the
    raw observations).

```{r}
t.test(PlantGrowth[PlantGrowth$group == "ctrl", 1],
       PlantGrowth[PlantGrowth$group != "ctrl", 1],
       alternative="less",
       var.equal = TRUE)
# The p.value ~0.41 there is no evidence to reject H0 as it is bigger than the 
#alpha 0.05. There is insufficient statistical evidence that the trtmean is 
# greater than the ctrl mean
```

As discussed, there is a rule of thumb for deciding whether to use a
pooled estimate of the variance in an unpaired t-test. e. Your task is
to write a wrapper function that calls t.test after deciding whether it
should be executed with var.equal=FALSE according to the rule of thumb.
Use the following guidelines: -- Your function should take four defined
arguments: x and y with no defaults, to be treated in the same way as
the same arguments in t.test; and var.equal and paired, with defaults
that are the same as the defaults of t.test. -- An ellipsis (Section
9.2.5) should be included to represent any additional arguments to be
passed to t.test. -- Upon execution, the function should determine
whether paired=FALSE. \* If paired is TRUE, then there is no need to
proceed with the check of a pooled variance. \* If paired is FALSE, then
the function should determine the value for var.equal automatically by
using the rule of thumb. -- If the value of var.equal was set
automatically, you can assume it will override any value of this
argument initially supplied by the user. -- Then, call t.test
appropriately.

```{r}
my.t.test <- function(x,
                      y,
                      var.equal = FALSE,
                      paired = FALSE,
                      ...) {
      if (!paired) {
            var.equal <- max(c(sd(x), sd(y))) /
                  min(c(sd(x), sd(y))) < 2
      }
      return(t.test(
            x = x,
            y = y,
            var.equal = var.equal,
            paired = paired,
            ...
      ))
}
my.t.test2 <- function(x,
                      y,
                      var.equal = FALSE,
                      paired = FALSE,
                      ...) {
      if (paired) {
            var.equal <- max(c(sd(x), sd(y))) /
                  min(c(sd(x), sd(y))) < 2
      }
      return(t.test(
            x = x,
            y = y,
            var.equal = var.equal,
            paired = paired,
            ...
      ))
}

```

f.  Try your new function on all three examples in the text of Section
    18.2.2, ensuring you reach identical results.

```{r}
#i 
snacks <- c(87.7,80.01,77.28,78.76,81.52,74.2,80.71,79.5,77.87,81.94,80.7,82.32,
            75.78,80.19,83.91,79.4,77.52,77.62,81.4,74.89,82.95,73.59,77.92,77.18,
            79.83,81.23,79.28,78.44,79.01,80.47,76.23,78.89,77.14,69.94,78.54,79.7,
            82.45,77.29,75.52,77.21,75.99,81.94,80.41,77.7)
snacks2 <- c(80.22,79.73,81.1,78.76,82.03,81.66,80.97,81.32,80.12,78.98,79.21,
             81.48,79.86,81.06,77.96,80.73,80.34,80.01,81.82,79.3,79.08,79.47,
             78.98,80.87,82.24,77.22,80.03,79.2,80.95,79.17,81)
my.t.test(snacks2,snacks,alternative="greater")


#ii
men <- c(102,87,101,96,107,101,91,85,108,67,85,82)
women <- c(73,81,111,109,143,95,92,120,93,89,119,79,90,126,62,92,77,106,105,111)

my.t.test(men,women,alternative="two.sided")

#iii
rate.before <- c(52,66,89,87,89,72,66,65,49,62,70,52,75,63,65,61) 
rate.after <- c(51,66,71,73,70,68,60,51,40,57,65,53,64,56,60,59) 

my.t.test(rate.after,rate.before,alternative="less",paired=TRUE)
```

Exercise 18.3

An advertisement for a skin cream claims nine out of ten women who use
it would recommend it to a friend. A skeptical salesperson in a
department store believes the true proportion of women users who'd
recommend it, , is much smaller than 0:9. She follows up with 89 random
customers who had purchased the skin cream and asks if they would
recommend it to others, to which 71 answer yes.

a.  Set up an appropriate pair of hypotheses for this test and determine
    whether it will be valid to carry out using the normal distribution.

```{r}
#Hypotheses: H0: p=0.9 // HA: p<0.9
p<-0.9
n<-89
x<-71
p.hat<-x/n
n*p.hat>5&n*(1-p.hat)>5
#as both are true, we can accept the to carry out with a normal distribution
```

b.  Compute the test statistic and the p-value and state your conclusion
    for the test using a significance level of  = 0:1.

```{r}
z.test<-(p.hat-p)/(sqrt(p*(1-p)/n))
pnorm(z.test)
#the P-value is too small. with 99% of confidence, there is  enought statistical
#evidenve to reject H0 in favor of HA.
```

c.  Using your estimated sample proportion, construct a two-sided 90
    percent confidence interval for the true proportion of women who
    would recommend the skin cream.

```{r}
p.hat+c(-1,1)*qnorm(0.99)*sqrt(p*(1-p)/n)
#this result corroborates the answer in "B" as it excludes the claimmed mean of 
#0.9
```

The political leaders of a particular country are curious as to the
proportion of citizens in two of its states that support the
decriminalization of marijuana. A small pilot survey taken by officials
reveals that 97 out of 445 randomly sampled voting-age citizens residing
in state 1 support the decriminalization and that 90 out of 419
votingage citizens residing in state 2 support the same notion.

d.  Letting 1 denote the true proportion of citizens in support of
    decriminalization in state 1, and 2 the same measure in state 2,
    conduct and conclude a hypothesis test under a significance level of
     = 0:05 with reference to the following hypotheses:

```{r}
x1<-97
n1<-445
x2<-90
n2<-419

p.hat1<-x1/n1
p.hat2<-x2/n2

#d
alpha<-0.05
#Hypotheses:: H0: p2-p1 =0 // HA:p2-p1 != 0
prop.test(c(x2,x1),
          c(n2,n1),
          alternative = "two.sided",
          conf.level = 1-alpha,
          correct = FALSE)
#the P-value is a big one. There is not enought statistical evidence that 
#the probability is different in each country. Pooled Variance
p<-(x1+x2)/(n1+n2)
#testing with the manual calculation
z<-(p.hat2-p.hat1)/(sqrt(p*(1-p)*(1/n1+1/n2)))
pval<-2*pnorm(z)
```

e.  Compute and interpret a corresponding CI. Though there is standard,
    ready-to-use R functionality for the t-test, at the time of this
    writing, there is no similar function for the Z-test (in other
    words, the normal-based test of proportions described here) except
    in contributed packages.

```{r}
# statistic+critical Value*SE
se<-sqrt(p*(1-p)*(1/n1+1/n2))
(p.hat2-p.hat1)+c(-1,1)*qnorm(1-alpha/2)*se
#the CI includes the value 0 reassuring that both states might have the same 
# proportion.
```

f.  Your task is to write a relatively simple R function, Z.test, that
    can perform a one- or two-sample Z-test, using the following
    guidelines: -- The function should take the following arguments: p1
    and n1 (no default) to pose as the estimated proportion and sample
    size; p2 and n2 (both defaulting to NULL) that contain the second
    sample proportion and sample size in the event of a two-sample test;
    p0 (no default) as the null value; and alternative (default
    "two.sided") and conf.level (default 0.95), to be used in the same
    way as in t.test. -- When conducting a two-sample test, it should be
    p1 that is tested as being smaller or larger than p2 when
    alternative="less" or alternative="greater", the same as in the use
    of x and y in t.test. -- The function should perform a one-sample
    Z-test using p1, n1, and p0 if either p2 or n2 (or both) is NULL.

-- The function should contain a check for the rule of thumb to ensure
the validity of the normal distribution in both one- and two-sample
settings. If this is violated, the function should still complete but
should issue an appropriate warning message (see Section 12.1.1). -- All
that need be returned is a list containing the members Z (test
statistic), P (appropriate p-value---this can be determined by
alternative; for a two-sided test, determining whether Z is positive or
not can help), and CI (two-sided CI with respect to conf.level).

```{r}
z.test <- function(p1,
                   n1,
                   p2 = NULL,
                   n2 = NULL,
                   p0,
                   alternative = "two.sided",
                   conf.level = 0.95) {
      
      if(is.null(p2)|is.null(n2)){
            cat("One sample Z-Test\n")
            if(p1*n1||n1*(1-p1)){
                  warning("normal distribution may not be valid")
            }
            z<-(p1-p0)/sqrt(p0*(1-p0)/n1)
            CI<-(p1)+c(-1,1)*qnorm(1-conf.level/2)*sqrt(p0*(1-p0)/n1)
      }else{
            cat("two-sample test")
            if(p1*n1<5||n1*(1-p1)<5||n2*p2<5||n2*(1-p2)<5){
                  warning("Normal distribution might not be accurate as rule 
                          of thumb")
            }
            p.star(((n1*p1)+(n2*p2))/n1+n2)
            z<-(p1-p2-p0)/sqrt(p.star*(1-p.star)*(1/n1+1/n2))
            CI<-(p1-p1)+c(-1,1)*qnorm(1-conf.level/2)*
                  sqrt(p.star*(1-p.star)*(1/n1+1/n2))
            }
      #defining p-value
      p<-pnorm(z)
      if(alternative=="greater"){
            p<-1-p
      }
      if(alternative=="two.sided"){
            if(z>0){
                  p<-2*(1-p)
            }else{
                  p<-2*p
            }
      }
      return(list(Z=z,P=p,CI=CI))
}

sick <- c(0,0,1,1,0,0,0,0,0,1,0,0,0,0,0,0,0,0,1,0,0,0,1,1,1,0,0,0,1)
n.sick<-length(sick)
p.sick<-sum(sick)/n.sick
```

g.  Replicate the two examples in the text of Sections 18.3.1 and 18.3.2
    using Z.test; ensure you reach identical results.
h.  Call Z.test(p1=0.11,n1=10,p0=0.1) to try your warning message in the
    one-sample setting.

Exercise 18.4

HairEyeColor is a ready-to-use data set in R that you haven't yet come
across. This 4 \* 4 \* 2 array provides frequencies of hair and eye
colors of 592 statistics students, split by sex (Snee, 1974).

a.  Perform and interpret, at a significance level of  = 0:01, a
    chisquared test of independence for hair against eye color for all
    students, regardless of their sex.

```{r}
data(HairEyeColor)
total.hec<-HairEyeColor[,,1]+HairEyeColor[,,2]

chisq.test(total.hec)
#H0:: Variables A and B are Independent // Ha: Variables are not independent
# We have a very small p.number. A strong statistical evidence to reject H0
```

In Exercise 8.1 on page 161, you accessed the Duncan data set of the
contributed package car, which contains markers of job prestige
collected in 1950. Install the package if you haven't already and load
the data frame. b. The first column of Duncan is the variable type,
recording the type of job as a factor with three levels: prof
(professional or managerial), bc (blue collar), and wc (white collar).
Construct appropriate hypotheses and perform a chi-squared GOF test to
determine whether the three job types are equally represented in the
data set. i. Interpret the resulting p-value with respect to a
significance level of  = 0:05.

ii. What conclusion would you reach if you used a significance level of
     = 0:01?

```{r}
library("car")
Duncan[,1]

c.prof<-length(Duncan[Duncan[,1]=="prof",1])
c.bc<-length(Duncan[Duncan[,1]=="bc",1])
c.wc<-length(Duncan[Duncan[,1]=="wc",1])


Duncan.tab<-table(Duncan[,1])

#Hypotheses: H0: p1=p2=p3=1 //Ha: h0 is incorrect

chisq.test(Duncan.tab)

#the P-value is lesser than the alpha=0.05, there is enough evidence to believe 
# the Variables are dependent, following Ha

#B
#with alpha=0.01, the p.value becomes greater than it, we tend with H0, the 
#variables are independent. 
```

Exercise 18.5

a.  Write a new version of typeI.tester called typeI.mean. The new
    function should be able to simulate the Type I error rate for tests
    of a single mean in any direction (in other words, oneor two-sided).
    The new function should take an additional argument, test, which
    takes a character string "less", "greater", or "two.sided" depending
    on the type of desired test. You can achieve this by modifying
    typeI.tester as follows: -- Instead of calculating and storing the
    p-values directly in the for loop, simply store the test statistic.

-- When the loop is complete, set up stacked if-else statements that
cater to each of the three types of test, calculating the p-value as
appropriate. -- For the two-sided test, remember that the p-value is
defined as twice the area "more extreme" than the null. Computationally,
this means you must use the upper-tail area if the test statistic is
positive and the lower-tail area otherwise. If this area is less than
half of  (since it is subsequently multiplied by 2 in a "real"
hypothesis test), then a rejection of the null should be flagged. -- If
the value of test is not one of the three possibilities, the function
should throw an appropriate error using stop. i. Experiment with your
function using the first example setting in the text with 0 = 0,  = 1,
n = 40, and  = 0:05. Call typeI.mean three times, using each of the
three possible options for test. You should find that all simulated
results sit close to 0:05.

```{r}
typeI.mean<-function(mu0,sigma,n,alpha,ITERATIONS=10000,test="less"){
      
      t.stat<-rep(NA,ITERATIONS)
      pvals<-rep(NA,ITERATIONS)
      for(i in 1:ITERATIONS){
            temporary.sample<-rnorm(n=n,mean=mu0,sd=sigma)
            temporary.mean<-mean(temporary.sample)
            temporary.sd<-sd(temporary.sample)
            t.stat[i]<-(temporary.mean-mu0)/(temporary.sd/sqrt(n))
      }
      if(test=="less"){
        pvals<-pt(t.stat,df=n-1)    
      }else 
            if(test=="greater"){
            pvals<-1-pt(t.stat,df=n-1)
      }else 
            if(test=="two.sided"){
            pvals[t.stat>=0]<-2*(1-pt(t.stat[t.stat>=0],df=n-1))
            pvals[t.stat<0]<-2*pt(t.stat[t.stat<0],df=n-1)
            
            }else {
            stop("Not a valid test input, please use \"less\",\"greater\", or 
\"two.sided\"")
      }
      return(mean(pvals<alpha))
}

# i
typeI.mean(0,1,40,0.05,test="less")
typeI.mean(0,1,40,0.05,test="greater")
typeI.mean(0,1,40,0.05,test="two.sided")

```

ii. Repeat (i) using the second example setting in the text with 0 =
    􀀀4,  = 0:3, n = 60, and  = 0:01. Again, you should find that all
    simulated results sit close to the value of .

```{r}
typeI.mean(-4,0.3,60,0.01,test="less")
typeI.mean(-4,0.3,60,0.01,test="greater")
typeI.mean(-4,0.3,60,0.01,test="two.sided")
```

b.  Modify typeII.tester in the same way as you did typeI.tester; call
    the new function typeII.mean. Simulate the Type II error rates for
    the following hypothesis tests. As per the text, assume A, , ,
    and n denote the true mean, standard deviation of raw observations,
    significance level, and sample size, respectively.

```{r}
typeII.mean<-function(mu0,muA,sigma,n,alpha,test="two.sided",ITERATIONS=10000){
      test.t<-rep(NA,ITERATIONS)
      for(i in 1:ITERATIONS){
            temporary.sample<-rnorm(n=n,mean=muA,sd=sigma)
            temporary.mean<-mean(temporary.sample)
            temporary.sd<-sd(temporary.sample)
            test.t[i]<-(temporary.mean-mu0)/(temporary.sd/sqrt(n))
      }
      pvals<-pt(test.t,df=n-1)
      if(test=="less"){
            return(mean(pvals>=alpha))
      }else if(test=="greater"){
            return(mean(1-pvals>=alpha))
      }else if(test=="two.sided"){
            result <- pvals
            result[test.t>0] <- 1-pvals[test.t>0]
            return(mean(result>=alpha/2))
      }else {
            stop("argument \"test\" is not valid. Please use \"less\", 
\"greater\" or \"two.sided\"")
      }
}

#i
typeII.mean(-3.2,-3.3,0.1,25,0.05,"two.sided")
#ii
typeII.mean(8994,5600,3888,9,0.01,"less")
#iii
typeII.mean(0.44,0.4,2.4,68,0.05,"greater")
```

Exercise 18.6

a.  For this exercise you'll need to have written typeII.mean from
    Exercise 18.5 (b). Using this function, modify power.tester so that
    a new function, power.mean, calls typeII.mean instead of calling
    typeII.tester.

```{r}
power.mean<-function(nvec,...){
      nlen<-length(nvec)
      result<-rep(NA,nlen)
      pbar<-txtProgressBar(min=0,max=nlen,style=3)
      for(i in 1:nlen){
            result[i]<-1-typeII.mean(n=nvec[i],...)
            setTxtProgressBar(pbar,i)
      }
      close(pbar)
      return(result)
}

#i
sample.sizes=5:100
power.mean(nvec=50,mu0=10,muA=10.5,sigma=0.9,alpha=0.01,test="two.sided")

#ii
power.mean(nvec=44,mu0=80,muA=78.5,sigma=3.1,alpha=0.05,test="two.sided")

power.mean(nvec=44,mu0=80,muA=78.5,sigma=3.1,alpha=0.01,test="two.sided")
```

b.  Staying with the snacks hypothesis test, using the sample.sizes
    vector from the text, determine the minimum sample size required for
    a statistically powerful test using both  = 0:05 and  = 0:01.
    Produce a plot showing the two power curves.

```{r}
snacks <- c(87.7,80.01,77.28,78.76,81.52,74.2,80.71,79.5,77.87,81.94,80.7,82.32,
            75.78,80.19,83.91,79.4,77.52,77.62,81.4,74.89,82.95,73.59,77.92,
            77.18,79.83,81.23,79.28,78.44,79.01,80.47,76.23,78.89,77.14,69.94,
            78.54,79.7,82.45,77.29,75.52,77.21,75.99,81.94,80.41,77.7)

sample.sizes<- 5:100




```

# chapter 19: ANALYSIS OF VARIANCE

Exercise 19.1

These figures provide the depths (in centimeters) at which important
archaeological finds were made at four sites in New Mexico (see Woosley
and Mcintyre, 1996). Store these data in your R workspace, with one
vector containing depth and the other vector containing the site of each
observation.

```{r}
info<-data.frame("dep"=c(93,120,65,105,115,82,99,87,100,90,78,95,93,88,110,85,
                         45,80,28,75,70,65,55,50,40,100,75,65,40,73,65,50,30,
                         45,50,45,55,96,58,95,90,65,80,85,95,82),
                 "site"=c(rep("Site i",15),
                          rep("Site ii",10),
                          rep("Site iii",12),
                          rep("Site iv",9)))
```

a.  Produce side-by-side boxplots of the depths split by group, and use
    additional points to mark the locations of the sample means.

```{r}
infomean<-tapply(info$dep,info$site,mean)

boxplot(info$dep~info$site,
        xlab="Site",
        ylab="Depth (cm)",
        main="Boxplot of Depth by cm by Site")
points(1:4,infomean,pch=4)
```

b.  Assuming independence, execute diagnostic checks for normality and
    equality of variances.

```{r}
infomeancen<-info$dep-infomean[as.numeric(info$site)]
qqnorm(infomeancen)
#The chart provides enought information to accept the normality of the data

info.sds<-tapply(info$dep,info$site,sd)
max(info.sds)/min(info.sds)
#to check the equality of variances we used the rule of thumb. as it is <2. We 
#accept it

```

c.  Perform and conclude a one-way ANOVA test for evidence of a
    difference between the means.

```{r}
infoaov<-aov(dep~site,info)
summary(infoaov)
#the p value is very small we have strong evidence to reject H0

dev.off()
```

In Section 14.4, you looked at the data set providing measurements on
petal and sepal sizes for three species of iris flowers. This is
available in R as iris.

d.  Based on diagnostic checks for normality and equality of variances,
    decide which of the four outcome measurements (sepal length/width
    and petal length/width) would be suitable for ANOVA (using the
    species as the group variable).

```{r}
sl.mean<-tapply(iris$Sepal.Length,iris$Species,mean) #mean
sl.meancen<-iris$Sepal.Length-sl.mean[as.numeric(iris$Species)] #mean centered
sl.sd<-tapply(iris$Sepal.Length,iris$Species,sd)

#checking variance
boxplot(iris$Sepal.Length~iris$Species)
      points(1:3,sl.mean,pch=4)

max(sl.sd)/min(sl.sd)
#the variance in Boxplot looks acceptable and the rule of thumb <2 confirms this

#checking normality
qqnorm(sl.meancen)
#the data gathered looks normal as well. we can use this information for an ANOVA.
#Sepal.width ----
sw.mean<-tapply(iris$Sepal.Width,iris$Species,mean) #mean
sw.meancen<-iris$Sepal.Width-sw.mean[as.numeric(iris$Species)] #mean centered
sw.sd<-tapply(iris$Sepal.Width,iris$Species,sd)

#checking variance
boxplot(iris$Sepal.Width~iris$Species)
      points(1:3,sl.mean,pch=4)

max(sw.sd)/min(sw.sd)
#the variance in Boxplot looks acceptable and the rule of thumb <2 confirms this

#checking normality
qqnorm(sw.meancen)
qqline(sw.meancen)
#the data gathered looks normal as well. we can use this information for an 
#ANOVA. This distributions looks mor normal then the previous
#Petal.Length ----
pl.mean<-tapply(iris$Petal.Length,iris$Species,mean) #mean
pl.meancen<-iris$Petal.Length-pl.mean[as.numeric(iris$Species)] #mean centered
pl.sd<-tapply(iris$Petal.Length,iris$Species,sd)

#checking variance
boxplot(iris$Petal.Length~iris$Species)
      points(1:3,sl.mean,pch=4)

max(pl.sd)/min(pl.sd)
#the variance in Boxplot does not look acceptable, and the rule of thumb is not 
#<2. this is not a good data to proceed with ANOVA.

#checking normality
qqnorm(pl.meancen)
qqline(pl.meancen)
#the data gathered looks normal as well. we can use this information for an 
#ANOVA. This distributions looks mor normal then the previous
#Petal.width----
pw.mean<-tapply(iris$Petal.Width,iris$Species,mean) #mean
pw.meancen<-iris$Petal.Width-pw.mean[as.numeric(iris$Species)] #mean centered
pw.sd<-tapply(iris$Petal.Width,iris$Species,sd)

#checking variance
boxplot(iris$Petal.Width~iris$Species)
      points(1:3,sl.mean,pch=4)

max(pw.sd)/min(pw.sd)
#the variance in Boxplot does not look acceptable
#and the rule of thumb >2 confirms this

#checking normality
qqnorm(pw.meancen)
qqline(pw.meancen)
#the data gathered looks normal as well. we can use this information for an 
#ANOVA. This distributions looks mor normal then the previous
#the best way to use a ANOVA is with the Sepal analysis
```

e.  Carry out one-way ANOVA for any suitable measurement variables.

```{r}
sl.aov<-aov(Sepal.Length~Species,iris)
summary(sl.aov)
#The pvalue is much less than the F value. We reject H0. 

sw.aov<-aov(Sepal.Width~Species,iris)
summary(sl.aov)
#The pvalue is much less than the F value. We reject H0. 
```

Exercise 19.2 Bring up the quakes data frame again, which describes the
locations, magnitudes, depths, and number of observation stations that
detected 1,000 seismic events off the coast of Fiji. a. Use cut (see
Section 4.3.3) to create a new factor vector defining the depths of each
event according to the following three categories: (0;200], (200;400],
and (400;680].

```{r}
depth.fac <- cut(quakes$depth,breaks=c(0,200,400,680))
table(depth.fac)
```

b.  Decide whether a one-way ANOVA or a Kruskal-Wallis test is more
    appropriate to use to compare the distributions of the number of
    detecting stations, split according to the three categories in (a).

```{r}
m <- tapply(quakes$stations,depth.fac,mean)
mc <- quakes$stations-m[as.numeric(depth.fac)]
hist(mc)
qqnorm(mc)
qqline(mc) # Data appear non-normal... Kruskal-Wallis preferred over 
# parametric one-way ANOVA
```

c.  Perform your choice of test in (b) (assume a  = 0:01 level of
    significance) and conclude.

```{r}
kruskal.test(quakes$stations~depth.fac) # P-value > 0.01 (just barely); 
#retain null. Minimal evidence, at best, of a difference in median 
#number of detecting stations according to depth.fac categories.
```

Load the MASS package with a call to library("MASS") if you haven't
already done so in the current R session. This package includes the
ready-to-use Cars93 data frame, which contains detailed data on 93 cars
for sale in the United States in 1993 (Lock, 1993; Venables and Ripley,
2002). d. Use aggregate to compute the mean length of the 93 cars, split
by two categorical variables: AirBags (type of airbags
available---levels are Driver & Passenger, Driver only, and None), and
Man.trans.avail (whether the car comes in a manual transmission---levels
are Yes and No).

```{r}
library("MASS")
cars.means <- aggregate(Cars93$Length,
                        by=list(Cars93$AirBags,
                                Cars93$Man.trans.avail),
                        FUN=mean)
cars.means
```

e.  Produce an interaction plot using the results in (d). Does there
    appear to be an interactive effect of AirBags with Man.trans.avail
    on the mean length of these cars (if you consider only these
    variables)?

```{r}
interaction.plot(x.factor=cars.means[,1],
                 trace.factor=cars.means[,2],
                 respons=cars.means$x,
                 trace.label="Manual avail.",
                 xlab="Airbags",
                 ylab="Mean length") 
# There is some visual indication of interactive behavior owing to the 
#non-parallel nature of the two lines; but rememeber there are no 
#measures of variability of the various group means displayed on this 
#plot...
```

f.  Fit a full two-way ANOVA model for the mean lengths according to the
    two grouping variables (assume satisfaction of all relevant
    assumptions). Is the interactive effect statistically significant?
    Is there evidence of any main effects?

```{r}
summary(aov(Length~AirBags+Man.trans.avail+AirBags:Man.trans.avail,data=Cars93)) # No formal statistical evidence of an interactive effect. Strong evidence of main effects, however -- both 'Airbags' and 'Man.trans.avail' appear to be related to car length.

```

Exercise 20.1 Continue to use the survey data frame from the package
MASS for the next few exercises. a. Using your fitted model of student
height on writing handspan, survfit, provide point estimates and 99
percent confidence intervals for the mean student height for handspans
of 12, 15:2, 17, and 19:9 cm.

```{r}
library("MASS")
survfit <- lm(Height~Wr.Hnd,data=survey)
#(a)
predict(survfit,newdata=data.frame(Wr.Hnd=c(12,15.2,17,19.9)),interval="confidence",level=0.99)
```

b.  In Section 20.1, you defined the object incomplete.obs, a numeric
    vector that provides the records of survey that were automatically
    removed from consideration when estimating the model parameters.
    Now, use the incomplete.obs vector along with survey and Equation
    (20.3) to calculate ˆ 0 and ˆ 1 in R. (Remember the functions
    mean, sd, and cor. Ensure your answers match the output from
    survfit.)

```{r}
incomplete.obs <- which(is.na(survey$Height)|is.na(survey$Wr.Hnd))
rho.xy <- cor(survey$Wr.Hnd,survey$Height,use="complete.obs")
b1 <- sd(survey$Height[-incomplete.obs])/sd(survey$Wr.Hnd[-incomplete.obs])*rho.xy
b1
b0 <- mean(survey$Height[-incomplete.obs])-b1*mean(survey$Wr.Hnd[-incomplete.obs])
b0
```

c.  The survey data frame has a number of other variables present aside
    from Height and Wr.Hnd. For this exercise, the end aim is to fit a
    simple linear model to predict the mean student height, but this
    time from their pulse rate, given in Pulse (continue to assume the
    conditions listed in Section 20.2 are satisfied).

d.  Fit the regression model and produce a scatterplot with the fitted
    line superimposed upon the data. Make sure you can write down the
    fitted model equation and keep the plot open.

```{r}
plot(survey$Height~survey$Pulse,xlab="Pulse rate (bpm)",ylab="Height (cm)")
survfit <- lm(Height~Pulse,data=survey)
survfit # Model equation is  y = 177.86 - 0.072x
abline(survfit,lwd=2)
```

ii. Identify and interpret the point estimate of the slope, as well as
    the outcome of the test associated with the hypotheses H0 : 1 = 0;
    HA : 1 , 0. Also find a 90 percent CI for the slope parameter.

```{r}
summary(survfit) # For each additional bpm, the mean student height is estimated to decrease by 0.072cm; p-value for slope is 0.275. No evidence to reject H0. Insufficient evidence to conclude that pulse rate affects mean student height.
confint(survfit,level=0.9)
```

iii. Using your model, add lines for 90 percent confidence and
     prediction interval bands on the plot from (i) and add a legend to
     differentiate between the lines.

```{r}
xseq <- data.frame(Pulse=seq(30,110,length=100))
survfit.ci <- predict(survfit,newdata=xseq,interval="confidence",level=0.9)
survfit.pi <- predict(survfit,newdata=xseq,interval="prediction",level=0.9)

plot(survey$Height~survey$Pulse,xlab="Pulse rate (bpm)",ylab="Height (cm)")
abline(survfit,lwd=2)
lines(xseq[,1],survfit.ci[,2],lty=2)
lines(xseq[,1],survfit.ci[,3],lty=2)
lines(xseq[,1],survfit.pi[,2],lty=2,col="grey")
lines(xseq[,1],survfit.pi[,3],lty=2,col="grey")
legend("topleft",legend=c("90% CI","90% PI"),lty=2,col=c("black","grey"))
```

iv. Create an incomplete.obs vector for the current "height on pulse"
    data. Use that vector to calculate the sample mean of the height
    observations that were used for the model fitted in (i). Then add a
    perfectly horizontal line to the plot at this mean (use color or
    line type options to avoid confusion with the other lines present).
    What do you notice? Does the plot support your conclusions from
    (ii)?

```{r}
incomplete.obs <- which(is.na(survey$Height)|is.na(survey$Pulse))
#abline(h=mean(survey$Height[-incomplete.obs]),col=2,lty=3,lwd=3) # The line sits in the middle of the CI bands without breaching them. This supports the conclusion that pulse rate is not significantly related to mean student height.

```

Next, examine the help file for the mtcars data set, which you first saw
in Exercise 13.4 on page 287. For this exercise, the goal is to model
fuel efficiency, measured in miles per gallon (MPG), in terms of the
overall weight of the vehicle (in thousands of pounds). d. Plot the
data---mpg on the y-axis and wt on the x-axis.

```{r}
?mtcars
plot(mtcars$mpg~mtcars$wt,xlab="Weight (lbs/1000)",ylab="MPG")
```

e.  Fit the simple linear regression model. Add the fitted line to the
    plot from (d).

```{r}
carfit <- lm(mpg~wt,data=mtcars)
carfit
attach(mtcars)
plot(wt,mpg)
abline(carfit,lwd=2)
```

f.  Write down the regression equation and interpret the point estimate
    of the slope. Is the effect of wt on mean mpg estimated to be
    statistically significant?

```{r}
summary(carfit) # mean MPG = 37.28 - 5.34*weight # For each extra 1000lbs of weight, the mean MPG decreases by 5.34; p-value for slope is very small---result is statistically significant---strong evidence to suggest that mean MPG changes according to weight (for cars of this era).

```

g.  Produce a point estimate and associated 95 percent PI for a car that
    weighs 6;000 lbs. Do you trust the model to predict observations
    accurately for this value of the explanatory variable? Why or why
    not?

```{r}
predict(carfit,newdata=data.frame(wt=6),interval="prediction",level=0.95) # Predicting at 6000lbs seems untrustworthy. Extrapolation is far enough outside the range of the observed data such that the associated PI has a lower limit that is negative, which makes no sense in terms of the response variable of MPG.

```

Exercise 20.2 Continue using the survey data frame from the package MASS
for the next few exercises. a. The survey data set has a variable named
Exer, a factor with k = 3 levels describing the amount of physical
exercise time each student gets: none, some, or frequent. Obtain a count
of the number of students in each category and produce side-by-side
boxplots of student height split by exercise.

```{r}
library("MASS")
#(a)
table(survey$Exer)
boxplot(survey$Height~survey$Exer)
```

b.  Assuming independence of the observations and normality as usual,
    fit a linear regression model with height as the response variable
    and exercise as the explanatory variable (dummy coding). What's the
    default reference level of the predictor? Produce a model summary.

```{r}
survfit <- lm(Height~Exer,data=survey)
summary(survfit) # The reference level of the predictor defaults to 
#the first level of the factor, which in this case (as is the default in R) is 
#alphabetically arranged to be 'Freq'.

```

c.  Draw a conclusion based on the fitted model from (b)---does it
    appear that exercise frequency has any impact on mean height? What
    is the nature of the estimated effect?

**\*\* We observe that both of the levels for which we've obtained
coefficient estimates yield p-values that suggest evidence the
coefficients are different to zero. The coefficient corresponding to
'some' has the smallest p-value of the two additive dummy levels. The
negative point estimates of both estimates tell us that the model
predicts the effect on height of being in either the 'none' or the
'some' categories, when compared to the 'frequent' category, is one of a
decrease. In other words, it appears those who exercise less than
'frequent' are shorter on average. The shortest mean height is reserved
for those in the 'none' exercise category; the estimated coefficient
(-5.58) is more extreme than the coefficient for some' (-4.21). Overall
statistical significance of the predictor (in terms of the effect of
exercise on height) is supported by the global (omnibus F-test) P-value
of 0.0035.\*\***

d.  Predict the mean heights of one individual in each of the three
    exercise categories, accompanied by 95 percent prediction intervals.
```{r}
one.of.each <- factor(levels(survey$Exer))
one.of.each
predict(survfit,newdata=data.frame(Exer=one.of.each),interval="prediction")

```

e.  Do you arrive at the same result and interpretation for the
    height-by-exercise model if you construct an ANOVA table using aov?

```{r}
summary(aov(Height~Exer,data=survey)) # Same 'global' P-value as the lm model 
#summary. There is evidence to suggest mean student height differs according to 
#exercise frequency.
```


f.  Is there any change to the outcome of (e) if you alter the model so
    that the reference level of the exercise variable is "none"? Would
    you expect there to be? Now, turn back to the ready-to-use mtcars
    data set. One of the variables in this data frame is qsec, described
    as the time in seconds it takes to race a quarter mile; another is
    gear, the number of forward gears (cars in this data set have either
    3, 4, or 5 gears).
```{r}
ExerReordered <- relevel(survey$Exer,ref="None")
levels(ExerReordered)
summary(aov(Height~ExerReordered,data=survey)) 
# There is no change to the omnibus F-test if we reorder the the reference level
#of exercise to be 'none', and nor should we expect there to be. The global test
#of a difference between the means doesn't depend on what we set the baseline 
#value of the response to be.

```
    

g.  Using the vectors straight from the data frame, fit a simple linear
    regression model with qsec as the response variable and gear as the
    explanatory variable and interpret the model summary.
```{r}
carfit <- lm(qsec~gear,data=mtcars)
summary(carfit) # The effect of 'gear' when treated as a continuous variable is 
#interpreted as a decrease in quarter-mile time of around 0.5 seconds, on 
#average, for each additional forward gear. However, there is no evidence that 
#this effect is different to zero, with a high P-value of 0.243 (and a similar 
#global p-value)
```

h.  Explicitly convert gear to a factor vector and refit the model.
    Compare the model summary with that from (g). What do you find?
```{r}
#(h)
carfit2 <- lm(qsec~factor(gear),data=mtcars)
summary(carfit2) # The effect of 'gear' when treated as a categorical variable 
#now appears to be statistically significant. Having 4 gears (as opposed to the 
#reference level of 3) seems to increase the mean quarter-mile time by 1.27 
#seconds, having 5 gears appears to decrease the mean quarter-mile time by 2.05 
#seconds. The global (omnibus F-test) p-value is also quite small, yielding 
#evidence in support of an effect of 'gear' on 'qsec'.
```

i.  Explain, with the aid of a relevant plot in the same style as the
    right image of Figure 20-6, why you think there is a difference
    between the two models (g) and (h).
```{r}
plot(mtcars$qsec~mtcars$gear,xlab="No. of Gears",ylab="Quarter-mile Time")
abline(carfit,lwd=2) # The plot indicates clearly that the difference between 
#the two models is due to the fact that the relationship cannot be well 
#explained by a continuous straight line. The first model therefore is incapable
#of realistically capturing the effect of changing categories in 'gear' on 
#'qsec'.
```





# chapter 21 :MULTIPLE LINEAR REGRESSION



